---
title: "Determining Optimal Simulation Repetition for Forest Stand Modeling: A Replication Analysis of the ForCEEPS Forest Growth Model"
author: "Clémentine de Montgolfier"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r options, include=FALSE}
# set working dir for all chunks to ..
knitr::opts_knit$set(root.dir = "../../")
# Set global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r data, message=FALSE, warning=FALSE}
load("data/forceeps_output/repetition_mean.RData") # all_data
```

# Abstract

Forest growth modeling is essential for sustainable forest management and climate change mitigation strategies. However, determining the optimal number of simulation repetitions required to achieve statistically reliable estimates is challenge in computational forest ecology. This study addresses this question for the ForCEEPS forest growth model using data from the Retz forest in France.

We conducted a comprehensive replication analysis using 5,000 individual ForCEEPS simulations to determine the minimum number of simulations (n) required for stable estimates of key forest metrics including basal area, volume, biomass, productivity, and mortality. Our methodology employed a two-stage approach: first, we assessed the convergence of the coefficient of variation (CV) across different numbers ForCEEPS dynamics (m); second, we analyzed how CV and standardized error decrease with increasing simulation numbers (n).

Results demonstrate that 10 repetitions (m = 10) are sufficient for coefficient of variation convergence across all tested forest metrics. For simulation numbers, our analysis reveals that 10 simulations (n = 10) provide adequate precision for most forest management applications. While 50 simulations offer superior precision, the marginal improvement does not justify the increased computational cost for our applications.

# Introduction

## Study Objectives

This study aims to establish the minimum number of simulations for the ForCEEPS forest model in order to make conclusions that are not dependant on the stochasticity of the model.
In order to do that we are going to calculate the divergence between dynamics. A dynamic being the mean of n simulation repetitions. But for this we also need to know between how many dynamics the coefficient of variation is stable. So we are going to determine : 

1. **Number of(m)**: Number of dynamics between which the coefficient of variation needs to be calculated.
2. **Simulation (n)**: How many individual simulations are needed to achieve acceptable precision in forest metric estimates?

## Study Site and Model Description

### Study Area

The analysis focuses on the Retz forest (patch RETZ_00137_02). See RETZ inventory csv for more details.

### ForCEEPS Model

ForCEEPS is a spatially explicit, individual-based forest model that simulates tree growth, competition, mortality, and regeneration processes. The model operates at the individual tree level and incorporates climate variables, soil characteristics, and species-specific physiological parameters. Some output variables include:

- Total basal area (m2/ha)
- Total volume (m3/ha)  
- Total biomass (t/ha)
- Annual productivity (m2/ha/year)
- Tree mortality (individuals/ha/year)

# Methods

## Simulation Framework

### Data Generation

We conducted 5,000 independent ForCEEPS simulations for the Retz forest patch, each spanning 80 years (2020-2100) with annual time steps. All simulations used identical initial forest conditions but different random seeds to capture stochastic variability in growth processes. The simulation protocol included:

- **Patch area**: 1,000 hectares
- **Simulation period**: 80 years (2020-2100)
- **Climate scenario**: Current climate conditions (retz_act.climate)
- **Management scenario**: 80_3_1_25_FSyl-80 (no management)
- **Species composition**: 9 potential species based on local forest inventory data

## Statistical Metrics

### Coefficient of Variation (CV)

The coefficient of variation was calculated as:

$$CV = \frac{\sigma}{\mu}$$

where sigma is the standard deviation and mu is the mean of the forest metric across repetitions.

### Standardized Error

The standardized error was calculated as:

$$SE = \frac{1.833 \times \sigma}{\sqrt{n}}$$

where 1.833 represents the 90% confidence interval multiplier for a t-distribution, sigma is the standard deviation, and n is the number of simulations.

### Convergence Criteria

We used the following convergence criteria:
- **Error thresholds**: Standardized error < 5% and < 10% of mean values

# Results

## Repetition Optimization Analysis (Stage 1)

The first stage of our analysis focused on determining the minimum number of dynamics (m) required for stable coefficient of variation estimates. We tested for 3 different dynamics (dynamics which are the mean of n = 2, 10, 50 forceeps simulations) for three key forest metrics.

### Analytical Approach

For each combination of simulation number (n) and repetition number (m), we:

1. Randomly sampled n simulations from our 5,000-simulation pool
2. Calculated mean values for each forest metric on the n simulations to get one dynamic
3. Repeated this process m times to generate m dynamics
4. Computed the coefficient of variation across these m dynamics
5. Assessed CV stability as m increased

```{r repetition-analysis-function}
m_analyse <- function(all_data, nb_simul = 10, variable = TotalBasalArea.m2.ha) {
  library(dplyr)
  library(ggplot2)
  
  # Create empty data frame to store final results
  df <- data.frame()
  
  # List of all available simulation IDs
  simulation <- 1:5000
  
  # Perform repetitions of random sampling
  for (rep in 1:round(5000 / nb_simul, 0)) {
    
    # Initialize a temporary data frame to hold results for this repetition
    oneRep <- data.frame()
    
    # Randomly sample 'nb_simul' unique simulation IDs without replacement
    l <- sample(simulation, nb_simul, replace = FALSE)
    
    # Filter 'all_data' to include only rows with the sampled simulations
    oneRep <- all_data[all_data$simulation %in% l, ] %>%
      mutate(n = nb_simul) %>%
      bind_rows(oneRep)
    
    # Remove used simulation IDs from the pool
    simulation <- simulation[!simulation %in% l]
    
    # Add repetition number and bind to the global results data frame
    df <- oneRep %>%
      mutate(rep = rep) %>%
      bind_rows(df)
  }
  
  # Group repetitions into different sizes
  df <- df %>%
    mutate(m = case_when(
      rep <= 5 ~ 5,
      rep <= 15 ~ 10,
      rep <= 65 ~ 50,
      rep <= 165 ~ 100,
      rep <= 315 ~ 150,
      rep <= 515 ~ 200,
      rep <= 765 ~ 250,
      rep <= 1065 ~ 300,
      rep <= 1415 ~ 350,
      rep <= 1815 ~ 400,
      TRUE ~ NA
    )) %>%
    filter(!is.na(m), rep < round(5000 / nb_simul, 0))
  
  # Compute mean of selected variable for each (rep, m, n, date)
  df_mean <- df %>%
    group_by(rep, m, n, date) %>%
    summarise(mean_val = mean({{ variable }}, na.rm = TRUE), .groups = 'drop')
  
  # Plot the mean trajectories
  p1 <- df_mean %>%
    ggplot(aes(x = date, y = mean_val, color = factor(rep))) +
    geom_line(alpha = 0.7) +
    facet_wrap(~m, labeller = label_both) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
      title = paste("Mean Trajectories for", deparse(substitute(variable))),
      subtitle = paste("Simulation group size (n) =", nb_simul),
      x = "Year",
      y = paste(deparse(substitute(variable)), "(units)"),
      caption = "Each line represents one repetition; facets show different repetition group sizes (m)"
    )
  
  # Calculate coefficient of variation (CV) and error
  df_cv <- df_mean %>%
    group_by(m, date) %>%
    summarise(
      cv = sd(mean_val, na.rm = TRUE) / mean(mean_val, na.rm = TRUE),
      erreur = 1.833 * sd(mean_val, na.rm = TRUE) / sqrt(n()),
      .groups = 'drop'
    ) %>%
    group_by(m) %>%
    summarise(
      cv_max = max(cv, na.rm = TRUE), 
      erreur_max = max(erreur, na.rm = TRUE), 
      cv_mean = mean(cv, na.rm = TRUE),
      erreur_mean = mean(erreur, na.rm = TRUE),
      .groups = 'drop'
    )
  
  p2 <- df_cv %>%
    ggplot(aes(x = m, y = cv_max)) +
    geom_line(color = "blue", size = 1.2) +
    geom_point(color = "blue", size = 2) +
    theme_minimal() +
    labs(
      title = "Coefficient of Variation Convergence",
      subtitle = paste("Maximum CV across time for n =", nb_simul, "simulations"),
      x = "Number of Repetitions (m)",
      y = "Maximum Coefficient of Variation",
      caption = "Lower values indicate better precision and stability"
    ) +
    theme(plot.title = element_text(size = 12, face = "bold"))
  
  return(list(
    plot_mean = p1,
    cv_table = df_cv,
    plot_cv = p2
  ))
}
```

### Results for Total Basal Area

#### N= 2 dynamics

```{r basal-area-2sim, fig.cap="Coefficient of variation analysis for total basal area with 2 simulations per dynamic. The left panel shows variation coefficient evolution accross dynamic numbers. The right panel demonstrates CV and errors as table."}
# Run
results <- m_analyse(all_data, nb_simul = 2, variable = totalBasalArea.m2.ha)

# Create summary table
cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 2)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

# Convert table to tableGrob
table_grob <- tableGrob(
  cv_summary,
  rows = NULL,
  theme = ttheme_default(
    core = list(fg_params = list(cex = 0.9)),
    colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
  )
)

# Display side by side
layout <- results$plot_cv + ylim(0, 0.2) + 
  theme(plot.title = element_text(size = 10)) + 
  table_grob
layout
```

**Analysis**: With only 2 simulations per dynamic, the coefficient of variation remains relatively high (approximately 0.10) and requires approximately 50 dynamics for stabilization. The standard error is substantial (4 m2/ha), indicating significant uncertainty in estimates.

#### N= 10 dynamics

```{r basal-area-10sim, fig.cap="Coefficient of variation analysis for total basal area with 10 simulations per dynamic. CV values stabilize around 0.05 with good convergence achieved at 10 dynamics, representing an optimal balance between precision and computational efficiency."}
# Run the simulation for 10 repetitions
results <- m_analyse(all_data, nb_simul = 10, variable = totalBasalArea.m2.ha)

# Create summary table
cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 2)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

table_grob <- tableGrob(
  cv_summary,
  rows = NULL,
  theme = ttheme_default(
    core = list(fg_params = list(cex = 0.9)),
    colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
  )
)

layout <- results$plot_cv + ylim(0, 0.2) + 
  theme(plot.title = element_text(size = 10)) + 
  table_grob
layout
```

**Analysis**: With 10 simulations per dynamic, the coefficient of variation decreases substantially to approximately 0.05 and achieves stable convergence at just 5-10 dynamics. The standard error reduces to 1 m2/ha, representing a significant improvement in precision.

#### N= 50 dynamics

```{r basal-area-50sim, fig.cap="Coefficient of variation analysis for total basal area with 50 simulations per dynamic. CV values are very low (approximately 0.02) and demonstrate excellent stability from the first dynamic group, indicating high precision with minimal uncertainty."}
# Run the simulation for 50 repetitions
results <- m_analyse(all_data, nb_simul = 50, variable = totalBasalArea.m2.ha)

cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 2)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

table_grob <- tableGrob(
  cv_summary,
  rows = NULL,
  theme = ttheme_default(
    core = list(fg_params = list(cex = 0.9)),
    colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
  )
)

layout <- results$plot_cv + ylim(0, 0.05) + 
  theme(plot.title = element_text(size = 10)) + 
  table_grob
layout
```

**Analysis**: With 50 simulations per dynamic, the coefficient of variation is extremely low (approximately 0.02) and achieves immediate stability from the first dynamic group.

### Results for Total Volume

#### N= 2 dynamics

```{r volume-2sim, fig.cap="Coefficient of variation analysis for total volume with 2 simulations per dynamic. Similar patterns to basal area are observed, with CV values requiring substantial dynamics for convergence."}
results <- m_analyse(all_data, nb_simul = 2, variable = totalVolume.m3.ha)

cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 1)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

table_grob <- tableGrob(cv_summary, rows = NULL, 
                       theme = ttheme_default(
                         core = list(fg_params = list(cex = 0.9)),
                         colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
                       ))

layout <- results$plot_cv + ylim(0, 0.2) + table_grob
layout
```

**Analysis**: Volume results demonstrate similar patterns to basal area, with high CV values (=0.12) requiring 50+ dynamics for stabilization when using only 2 simulations per dynamic.

#### N= 10 dynamics

```{r volume-10sim, fig.cap="Coefficient of variation analysis for total volume with 10 simulations per dynamic. Convergence patterns mirror those observed for basal area, confirming the robustness of our findings across forest metrics."}
results <- m_analyse(all_data, nb_simul = 10, variable = totalVolume.m3.ha)

cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 1)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

table_grob <- tableGrob(cv_summary, rows = NULL,
                       theme = ttheme_default(
                         core = list(fg_params = list(cex = 0.9)),
                         colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
                       ))

layout <- results$plot_cv + ylim(0, 0.2) + table_grob
layout
```

**Analysis**: Volume metrics with 10 simulations per dynamic show consistent improvement over the n=2 scenario, with CV values stabilizing around 0.05 and convergence achieved within 10 dynamics.

### Results for Total Biomass

#### N= 2 dynamics

```{r biomass-2sim, fig.cap="Coefficient of variation analysis for total biomass with 2 simulations per dynamic. Biomass estimates show similar convergence patterns to other forest metrics, validating our methodological approach."}
results <- m_analyse(all_data, nb_simul = 2, variable = totalBiomass.t.ha)

cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 1)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

table_grob <- tableGrob(cv_summary, rows = NULL,
                       theme = ttheme_default(
                         core = list(fg_params = list(cex = 0.9)),
                         colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
                       ))

layout <- results$plot_cv + ylim(0, 0.2) + table_grob
layout
```

**Analysis**: Biomass estimates with 2 simulations per dynamic demonstrate similar high variability patterns as observed for basal area and volume, requiring substantial dynamics for convergence.

#### N= 10 dynamics

```{r biomass-10sim, fig.cap="Coefficient of variation analysis for total biomass with 10 simulations per dynamic. Results consistently support the conclusion that 10 dynamics provide adequate convergence for forest modeling applications."}
results <- m_analyse(all_data, nb_simul = 10, variable = totalBiomass.t.ha)

cv_summary <- results$cv_table %>%
  select(m, cv_max, erreur_max) %>%
  mutate(
    cv_max = round(cv_max, 4),
    erreur_max = round(erreur_max, 1)
  ) %>%
  rename(
    "Repetitions (m)" = m,
    "Max CV" = cv_max,
    "Max Error" = erreur_max
  )

table_grob <- tableGrob(cv_summary, rows = NULL,
                       theme = ttheme_default(
                         core = list(fg_params = list(cex = 0.9)),
                         colhead = list(fg_params = list(cex = 0.9, fontface = "bold"))
                       ))

layout <- results$plot_cv + ylim(0, 0.2) + table_grob
layout
```

**Analysis**: Biomass estimates with 10 simulations per dynamic consistently demonstrate optimal convergence behavior, supporting the general conclusion that 10 dynamics provide adequate precision for forest modeling applications. I skipped the 50 simulations analysis for biomass as the patterns are consistent with previous metrics.

### Stage 1 Conclusions

The analysis demonstrates that while 50 simulations provide superior precision from the outset, 10 repetitions achieve acceptable precision when using the mean of at least 10 simulations, establishing our framework for Stage 2 analysis.

## Simulation Optimization Analysis (Stage 2)

Having established that 10 repetitions provide optimal convergence (m = 10), we now focus on determining the minimum number of simulations (n) required to achieve acceptable precision levels. This analysis examines how coefficient of variation and standardized error decrease as simulation numbers increase from 1 to 150.

### Analytical Framework

Using m = 10 repetitions, we tested simulation group sizes ranging from 1 to 150 simulations. For each simulation group size, we:

1. Generated 10 independent repetitions, each containing n simulations
2. Calculated mean forest metrics for each repetition
3. Computed coefficient of variation and standardized error across repetitions
4. Assessed precision improvement with increasing simulation numbers

```{r simulation-optimization-function}
n_analyses <- function(all_data, variable = TotalBasalArea.m2.ha, 
                      groupes = c(1, 5, 10, 20, 30, 50, 75, 100, 150), 
                      nb_rep = 10) {
  library(dplyr)
  library(ggplot2)
  
  df <- data.frame()
  simulation <- 1:5000
  
  # Loop over repetitions
  for (m in 1:nb_rep) {
    oneRep <- data.frame()
    for (i in seq_along(groupes)) {
      l <- sample(simulation, groupes[i], replace = FALSE)
      oneRep <- all_data[all_data$simulation %in% l, ] %>%
        mutate(n = groupes[i]) %>%
        bind_rows(oneRep)
      simulation <- setdiff(simulation, l)
    }
    df <- oneRep %>%
      mutate(m = m) %>%
      bind_rows(df)
  }
  
  # Calculate means by (m, n, date)
  df_mean <- df %>%
    group_by(m, n, date) %>%
    summarise(mean_val = mean({{ variable }}, na.rm = TRUE), .groups = "drop")
  
  # Plot mean trajectories
  plot_mean <- df_mean %>%
    ggplot(aes(x = date, y = mean_val, color = factor(m))) +
    geom_line(alpha = 0.8) +
    facet_wrap(~n, labeller = label_both) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
      title = paste("Mean Trajectories for", deparse(substitute(variable))),
      subtitle = "Each line represents one repetition (m = 10 total)",
      x = "Year",
      y = paste(deparse(substitute(variable)), "(units)"),
      caption = "Facets show different simulation group sizes (n)"
    )
  
  # Calculate coefficient of variation and error
  df_cv <- df_mean %>%
    group_by(n, date) %>%
    summarise(
      cv = sd(mean_val, na.rm = TRUE) / mean(mean_val, na.rm = TRUE),
      error = 1.833 * sd(mean_val, na.rm = TRUE) / sqrt(n()),
      .groups = 'drop'
    ) %>%
    group_by(n) %>%
    summarise(
      cv_max = max(cv, na.rm = TRUE),
      error_max = max(error, na.rm = TRUE),
      cv_mean = mean(cv, na.rm = TRUE),
      error_mean = mean(error, na.rm = TRUE),
      .groups = 'drop'
    )
  
  # CV and Error plots
  plot_cv <- df_cv %>%
    ggplot(aes(x = n, y = cv_max)) +
    geom_line(color = "blue", size = 1.2) +
    geom_point(color = "blue", size = 2) +
    theme_minimal() +
    ylim(0, 0.2) +
    labs(
      title = "Coefficient of Variation vs. Simulation Number",
      x = "Number of Simulations (n)",
      y = "Maximum CV",
      caption = "Lower values indicate better precision"
    )
  
  plot_error <- df_cv %>%
    ggplot(aes(x = n, y = error_max)) +
    geom_line(color = "red", size = 1.2) +
    geom_point(color = "red", size = 2) +
    theme_minimal() +
    labs(
      title = "Standardized Error vs. Simulation Number",
      x = "Number of Simulations (n)",
      y = "Maximum Standardized Error",
      caption = "Error represents 90% confidence interval bounds"
    )
  
  return(list(
    plot_mean = plot_mean,
    plot_cv = plot_cv,
    plot_error = plot_error,
    summary_table = df_cv
  ))
}
```

### Results for Total Basal Area

```{r basal-area-simulation-analysis, fig.cap="Simulation optimization analysis for total basal area. Left panel shows mean trajectories across different simulation group sizes. Center panel demonstrates CV reduction with increasing simulation numbers. Right panel shows standardized error convergence, with significant improvements observed up to n=20."}
# Execute analysis for totalBasalArea.m2.ha
results <- n_analyses(all_data, variable = totalBasalArea.m2.ha)

# Arrange plots side by side
grid.arrange(results$plot_mean, results$plot_cv, results$plot_error, ncol = 3)
```

**Analysis**: Basal area estimates show rapid precision improvement up to n=10-20 simulations, with diminishing returns beyond n=50. The standardized error decreases substantially from n=1 to n=20, with minimal additional improvement at higher simulation numbers.

### Results for Total Volume

```{r volume-simulation-analysis, fig.cap="Simulation optimization analysis for total volume. Volume estimates show similar convergence patterns to basal area, with optimal precision achieved around n=10-20 simulations."}
# Execute analysis for totalVolume.m3.ha
results <- n_analyses(all_data, variable = totalVolume.m3.ha)

# Arrange plots side by side
grid.arrange(results$plot_mean, results$plot_cv, results$plot_error, ncol = 3)
```

**Analysis**: Volume metrics demonstrate consistent convergence patterns with basal area estimates, confirming the robustness of our simulation optimization approach across different forest metrics.

### Results for Total Biomass

```{r biomass-simulation-analysis, fig.cap="Simulation optimization analysis for total biomass. Biomass estimates demonstrate consistent convergence behavior, supporting the robustness of our optimization approach across multiple forest metrics."}
# Execute analysis for totalBiomass.t.ha
results <- n_analyses(all_data, variable = totalBiomass.t.ha)

# Arrange plots side by side
grid.arrange(results$plot_mean, results$plot_cv, results$plot_error, ncol = 3)
```

**Analysis**: Biomass estimates demonstrate consistent convergence behavior across all simulation numbers, further validating the robustness of our optimization approach across multiple forest metrics.

## Analysis of Productivity Metrics

Working on productivity analysis to:
- Obtain more temporally stable values (less variation, making error less date-dependent)
- Reduce dependence on initial values (strong divergence in early years affects other accumulated metrics)

### Species-Level Productivity Analysis

```{r productivity-data-loading}
load("data/forceeps_output/repetition_productivityScene.RData") # global_df
```

### Species-Level Results

```{r species-productivity-analysis, message=FALSE, warning=FALSE, fig.cap="Species-level productivity analysis showing coefficient of variation and error patterns across different simulation group sizes for individual tree species."}
global_df_sp <- global_df %>%
  filter(speciesShortName != "Tot")

# Plot mean trajectories
plot_mean <- global_df_sp %>%
  group_by(date, speciesShortName, n, m) %>%
  summarise(mean_val = mean(mean_val, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = date, y = mean_val, color = factor(m))) +
  geom_line(alpha = 0.7) +
  facet_grid(speciesShortName ~ n, labeller = label_both) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 8)
  ) +
  labs(
    title = "Mean Productivity Trajectories by Species",
    subtitle = "Adult basal area production (m2/ha/year)",
    x = "Year",
    y = "Adult Productivity (m2/ha/year)",
    caption = "Rows = species, Columns = simulation group size (n)"
  )

# Calculate coefficient of variation and error
df_cv <- global_df_sp %>%
  group_by(n, date, speciesShortName) %>%
  summarise(
    sd_val = sd(mean_val, na.rm = TRUE),
    cv = sd(mean_val, na.rm = TRUE) / mean(mean_val, na.rm = TRUE),
    error = 1.833 * sd(mean_val, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  group_by(n) %>%
  summarise(
    sd_max = max(sd_val, na.rm = TRUE),
    cv_max = max(cv, na.rm = TRUE),
    error_max = max(error, na.rm = TRUE),
    .groups = 'drop'
  )

# CV and Error plots
plot_cv <- df_cv %>%
  ggplot(aes(x = n, y = cv_max)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(color = "blue", size = 2) +
  theme_minimal() +
  labs(
    title = "Maximum CV Across Species",
    x = "Number of Simulations (n)",
    y = "Maximum CV",
    caption = "CV calculated across species and time"
  )

plot_error <- df_cv %>%
  ggplot(aes(x = n, y = error_max)) +
  geom_line(color = "red", size = 1.2) +
  geom_point(color = "red", size = 2) +
  theme_minimal() +
  labs(
    title = "Maximum Standardized Error",
    x = "Number of Simulations (n)",
    y = "Maximum Error",
    caption = "Error represents 90% confidence bounds"
  )

plot_mean / (plot_cv + plot_error) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "Species-Level Productivity Analysis",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
  )
```

**Species-Level Analysis Notes**: Coefficient of variation is unreliable for some species because certain values are zero (making the mean null and CV infinite). Additionally, some species have very low productivity, resulting in very high coefficient of variation values.

### Total Productivity Analysis

```{r total-productivity-analysis, message=FALSE, warning=FALSE, fig.cap="Total productivity analysis across all species, showing convergence patterns for aggregated productivity metrics with precision thresholds at 5% and 10% error levels."}
global_df_tot <- global_df %>%
  filter(speciesShortName == "Tot")

# Calculate threshold values at 5% and 10% of mean
threshold_5pct <- 0.05 * mean(global_df_tot$mean_val, na.rm = TRUE)
threshold_10pct <- 0.1 * mean(global_df_tot$mean_val, na.rm = TRUE)

# Plot mean trajectories
plot_mean_tot <- global_df_tot %>%
  group_by(date, n, m) %>%
  summarise(mean_val = mean(mean_val, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = date, y = mean_val, color = factor(m))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~n, labeller = label_both) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Total Productivity Trajectories",
    subtitle = "Aggregate adult basal area production across all species",
    x = "Year",
    y = "Total Adult Productivity (m2/ha/year)",
    caption = "Each line represents one repetition (m = 10 total)"
  )

# Calculate coefficient of variation and error
df_cv_tot <- global_df_tot %>%
  group_by(n, date) %>%
  summarise(
    sd_val = sd(mean_val, na.rm = TRUE),
    cv = sd(mean_val, na.rm = TRUE) / mean(mean_val, na.rm = TRUE),
    error = 1.833 * sd(mean_val, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  group_by(n) %>%
  summarise(
    sd_max = max(sd_val, na.rm = TRUE),
    cv_max = max(cv, na.rm = TRUE),
    error_max = max(error, na.rm = TRUE),
    .groups = 'drop'
  )

# CV and Error plots with thresholds
plot_cv_tot <- df_cv_tot %>%
  ggplot(aes(x = n, y = cv_max)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(color = "blue", size = 2) +
  theme_minimal() +
  labs(
    title = "Coefficient of Variation (Total)",
    x = "Number of Simulations (n)",
    y = "Maximum CV"
  )

plot_error_tot <- df_cv_tot %>%
  ggplot(aes(x = n, y = error_max)) +
  geom_line(color = "red", size = 1.2) +
  geom_point(color = "red", size = 2) +
  theme_minimal() +
  labs(
    title = "Standardized Error (Total)",
    x = "Number of Simulations (n)",
    y = "Maximum Error"
  ) +
  geom_hline(yintercept = threshold_5pct, linetype = "dashed", color = "red", alpha = 0.7) +
  geom_hline(yintercept = threshold_10pct, linetype = "dashed", color = "blue", alpha = 0.7) +
  annotate("text", x = max(df_cv_tot$n) * 0.8, y = threshold_5pct, 
           label = "5% threshold", color = "red", vjust = -0.5, size = 3) +
  annotate("text", x = max(df_cv_tot$n) * 0.8, y = threshold_10pct, 
           label = "10% threshold", color = "blue", vjust = 1.5, size = 3)

# Find intersection values with thresholds
intersect_5pct <- df_cv_tot %>%
  filter(error_max <= threshold_5pct) %>%
  summarise(n_min = min(n, na.rm = TRUE)) %>%
  pull(n_min)

intersect_10pct <- df_cv_tot %>%
  filter(error_max <= threshold_10pct) %>%
  summarise(n_min = min(n, na.rm = TRUE)) %>%
  pull(n_min)

plot_mean_tot / (plot_cv_tot + plot_error_tot) +
  plot_layout(heights = c(2, 1)) +
  plot_annotation(
    title = "Total Productivity Analysis",
    subtitle = paste("Precision thresholds: n =", intersect_10pct, "for 10% error, n =", intersect_5pct, "for 5% error"),
    theme = theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
  )
```

**Threshold Analysis Results**: 
- For 5% error threshold: **50 simulations** required  
- For 10% error threshold: **10 simulations** required

### Mortality Analysis

Analyzing tree mortality (deadNumber_ha) from the productivity scene dataset.

```{r mortality-analysis, message=FALSE, warning=FALSE, fig.cap="Tree mortality analysis showing the number of simulations required to achieve stable estimates of forest mortality rates with defined precision thresholds."}
global_df_mortality <- global_df %>%
  filter(speciesShortName == "Tot")

# Calculate threshold values for mortality
threshold_mort_5pct <- 0.05 * mean(global_df_mortality$dead_total, na.rm = TRUE)
threshold_mort_10pct <- 0.1 * mean(global_df_mortality$dead_total, na.rm = TRUE)

# Plot mean mortality trajectories
plot_mean_mortality <- global_df_mortality %>%
  group_by(date, n, m) %>%
  summarise(dead_total = mean(dead_total, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = date, y = dead_total, color = factor(m))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~n, labeller = label_both) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Tree Mortality Trajectories",
    subtitle = "Dead trees per hectare across simulation scenarios",
    x = "Year",
    y = "Dead Trees (individuals/ha)",
    caption = "Each line represents one repetition (m = 10 total)"
  )

# Calculate coefficient of variation and error for mortality
df_cv_mortality <- global_df_mortality %>%
  group_by(n, date) %>%
  summarise(
    sd_val = sd(dead_total, na.rm = TRUE),
    cv = sd(dead_total, na.rm = TRUE) / mean(dead_total, na.rm = TRUE),
    error = 1.833 * sd(dead_total, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  group_by(n) %>%
  summarise(
    sd_max = max(sd_val, na.rm = TRUE),
    cv_max = max(cv, na.rm = TRUE),
    error_max = max(error, na.rm = TRUE),
    .groups = 'drop'
  )

# CV and Error plots for mortality
plot_cv_mortality <- df_cv_mortality %>%
  ggplot(aes(x = n, y = cv_max)) +
  geom_line(color = "darkgreen", size = 1.2) +
  geom_point(color = "darkgreen", size = 2) +
  theme_minimal() +
  labs(
    title = "CV for Tree Mortality",
    x = "Number of Simulations (n)",
    y = "Maximum CV"
  )

plot_error_mortality <- df_cv_mortality %>%
  ggplot(aes(x = n, y = error_max)) +
  geom_line(color = "darkred", size = 1.2) +
  geom_point(color = "darkred", size = 2) +
  theme_minimal() +
  labs(
    title = "Error for Tree Mortality",
    x = "Number of Simulations (n)",
    y = "Maximum Error"
  ) +
  geom_hline(yintercept = threshold_mort_5pct, linetype = "dashed", color = "red", alpha = 0.7) +
  geom_hline(yintercept = threshold_mort_10pct, linetype = "dashed", color = "blue", alpha = 0.7) +
  annotate("text", x = max(df_cv_mortality$n) * 0.8, y = threshold_mort_5pct, 
           label = "5% threshold", color = "red", vjust = -0.5, size = 3) +
  annotate("text", x = max(df_cv_mortality$n) * 0.8, y = threshold_mort_10pct, 
           label = "10% threshold", color = "blue", vjust = 1.5, size = 3)

# Find intersection values for mortality
intersect_mort_5pct <- df_cv_mortality %>%
  filter(error_max <= threshold_mort_5pct) %>%
  summarise(n_min = min(n, na.rm = TRUE)) %>%
  pull(n_min)

intersect_mort_10pct <- df_cv_mortality %>%
  filter(error_max <= threshold_mort_10pct) %>%
  summarise(n_min = min(n, na.rm = TRUE)) %>%
  pull(n_min)

plot_mean_mortality / (plot_cv_mortality + plot_error_mortality) +
  plot_layout(heights = c(2, 1)) +
  plot_annotation(
    title = "Tree Mortality Analysis",
    subtitle = paste("Precision requirements: n =", intersect_mort_10pct, "for 10% error, n =", intersect_mort_5pct, "for 5% error"),
    theme = theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
  )
```

**Mortality Analysis Results**:
- n = **10** for 10.5% error
- n = **20** for 10% error  
- n = **50** for 5% error

# Discussion

Our analysis supports different simulation parameters based on application requirements:

- 10% error tolerance : n = 10, m = 10
- 5% error tolerance : n = 50, m = 10

## Methodological Considerations

### Strengths

1. **Comprehensive scope**: Analysis covered multiple forest metrics across 80-year simulations
2. **Large sample size**: 5,000 independent simulations provide robust statistical foundation
3. **Systematic approach**: Two-stage optimization ensures both repetition and simulation parameters are optimized
4. **Practical focus**: Results directly applicable to operational forest modeling

### Limitations

1. **Single forest type**: Results based solely on one initial inventory (to test on different inventories would be a good idea)
2. **Fixed time horizon**: 80-year simulation period may not capture long-term dynamics
3. **Limited climate scenarios**: Analysis used current climate conditions only
4. **Management regime**: Analyses were done with no management.


## Other way of determination, discarded

Before this analyses we conducted an other one. We discarded this approach because we decided on using a more robust method with variation coefficient and more independency between groups of repetitions. Howevec I keep this one in the document for the sake of transparency, and in order to understand our process and choices.

```{r precision-repetition-analysis, message=FALSE, warning=FALSE}
# Create groups of repetitions
groupes <- c(1, 5, 10, 20, 30, 50, 75, 100, 150)

df <- data.frame()
simulation <- 1:5000

for(rep in 1:10){
    oneRep <- data.frame()
    for(i in 1:length(groupes)){
        l <- sample(simulation, groupes[i], replace = FALSE)
        oneRep <- all_data[all_data$simulation %in% l,] %>%
            mutate(n = groupes[i]) %>%
            bind_rows(oneRep)
        simulation <- simulation[!simulation %in% l]
    }
    df <- oneRep %>%
        mutate(rep = rep) %>%
        bind_rows(df)
}

# Calculate means for stability analysis
df_mean <- df %>%
    group_by(rep, n, date) %>%
    summarise(totalBasalArea = mean(totalBasalArea.m2.ha, na.rm = TRUE), .groups = 'drop')

# Calculate temporal stability (epsilon)
df_epsilon <- df_mean %>%
    arrange(rep, n, date) %>%
    group_by(rep, n) %>%
    summarise(
        epsilon = sum(abs(totalBasalArea - lag(totalBasalArea, default = first(totalBasalArea))), na.rm = TRUE) / 82,
        .groups = 'drop'
    )

# Create stability plot
plot_stability <- ggplot(df_epsilon) +
    geom_line(aes(x = n, y = epsilon, color = factor(rep)), alpha = 0.7) +
    geom_line(
        data = df_epsilon %>% group_by(n) %>% summarise(epsilon = mean(epsilon), .groups = 'drop'),
        aes(x = n, y = epsilon), 
        color = "black", 
        size = 1.5
    ) +
    geom_hline(yintercept = 0.2, linetype = "dashed", color = "red", alpha = 0.7) +
    theme_minimal() +
    labs(
        title = "Temporal Stability Analysis",
        subtitle = "Lower epsilon values indicate more stable trajectories",
        x = "Number of Simulations (n)",
        y = "Epsilon (Temporal Variation Index)",
        color = "Repetition",
        caption = "Red dashed line represents acceptable stability threshold (0.2)"
    ) +
    theme(legend.position = "none")

# Calculate coefficient of variation across repetitions
df_cv_analysis <- df_mean %>%
    group_by(n, date) %>%
    summarise(
        cv = sd(totalBasalArea, na.rm = TRUE) / mean(totalBasalArea, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(n) %>%
    summarise(
        cv_mean = mean(cv, na.rm = TRUE),
        .groups = 'drop'
    )

plot_cv_analysis <- ggplot(df_cv_analysis) +
    geom_line(aes(x = n, y = cv_mean), color = "blue", size = 1.5) +
    geom_point(aes(x = n, y = cv_mean), color = "blue", size = 2) +
    theme_minimal() +
    labs(
        title = "Mean Coefficient of Variation",
        subtitle = "CV averaged across the 80-year simulation period",
        x = "Number of Simulations (n)",
        y = "Mean CV",
        caption = "Lower values indicate better precision"
    )

grid.arrange(plot_stability, plot_cv_analysis, ncol = 2)
```