---
title: "Analysis of Inventory Generation Methods for Forest Simulation"
author: "Clémentine de Montgolfier"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r options, include=FALSE}
# set working dir for all chunks to ..
knitr::opts_knit$set(root.dir = "../../")
# Set global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# Abstract

This study tests different methods for generating initial forest inventories for the FORCEEPS simulation model. We compare the effects of various diameter distributions and species proportions on simulation outputs.

The initial data comes from the Retz forest, providing basal area, dominant diameter intervals, and species proportions rounded to 10%. This must be converted to a FORCEEPS inventory format for 1000 m² plots (max 3000 m²), raising several methodological questions:

- How to handle species representation when low percentages result in fractional trees (e.g., 2% beech in 20 trees = 0.4 trees)?
- Should we test variations around initial data given unknown uncertainties and their potential effects on forest dynamics?

This preliminary analysis explores these questions and provides a foundation for more comprehensive future research.

# Initial Inventory Analysis

```{r}
# Load required packages and data
load("data/forceeps_output/inventory.RData")
```

**Figure 1: Basal area by species in initial inventories**

This analysis uses a single inventory from the Retz forest. Different functions were used to create the inventories:

- **Diameter distribution methods**: max, min, mean, normal_6, normal_4, unif (representing different statistical approaches to diameter allocation)
- **Species proportion methods**: fixe (fixed) or aléatoire (random) (representing deterministic vs. stochastic species allocation)

```{r}
# Calculate basal area for each tree and create inventory identifiers
Inventories <- Inventories %>%
    mutate(inventaire_id = paste(distribution, species_proportion, repetition, sep = "_"),
        surf.ter = (pi * ((dbh/100)/2)^2)*10)

# Plot basal area by inventory and species
ggplot(Inventories, aes(x = inventaire_id, y = surf.ter, fill = factor(speciesId))) +
    geom_bar(stat = "identity", position = "stack") +
    labs(y = "Surface terrière", x = "Inventaire", fill = "Espèce") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Data Processing and Loading

```{r}
# Clean and rename FORCEEPS output files for analysis
# Process files from multiple simulation seeds for robust results
seeds = c(332, 124, 102, 895, 869, 777, 969, 449, 131, 704)
for (i in seeds){
    # Define folder containing the output files
    folder <- 
        paste0(base_path, "output-cmd_",
        as.character(i), ".txt/")  # Replace with your path

    # List all files in the folder
    files <- list.files(folder, full.names = TRUE)

    # Define patterns to clean from filenames
    prefix_pattern <- "^retz_act\\.climate_inventaires_"
    simul_pattern <- "_simulation_([1-9]|10|11)"
    inv_pattern <- "\\.inv"

    # Process and rename files for easier analysis
    for (file in files) {
      filename <- basename(file)
    
      # Apply filename cleaning
      new_filename <- filename
      new_filename <- sub(prefix_pattern, "", new_filename)
      new_filename <- sub(simul_pattern, "", new_filename)
      new_filename <- sub(inv_pattern, "_", new_filename)  # Replace .inv with _

      # Rename file if changes were made
      if (new_filename != filename) {
        old_path <- file.path(folder, filename)
        new_path <- file.path(folder, new_filename)

        file.rename(old_path, new_path)
      }
    }
}

```

```{r}

# Load and process FORCEEPS simulation data
seeds <- c(332, 124, 102, 895, 869, 777, 969, 449, 131, 704)

# Function to read and parse simulation output files
read_forceps_file <- function(seed) {
    folder <- paste0(base_path, "output-cmd_", seed, ".txt/")
    files <- list.files(folder, pattern = "_productivityScene\\.txt$", full.names = TRUE)
    
    if (length(files) == 0) return(NULL)
    
    # Process each file and extract metadata from filename
    map_dfr(files, function(f) {
        # Extract information from filename
        fname <- basename(f)
        # Expected pattern: {distribution}_{species_proportion}_{repetition}_productivityScene.txt
        m <- str_match(fname, "^(.+?)_([a-zA-Z0-9]+)_([0-9]+)_productivityScene\\.txt$")
        if (is.na(m[1,1])) return(NULL)
        distribution <- m[1,2]
        species_proportion <- m[1,3]
        repetition <- as.integer(m[1,4]) + 1 # +1 because files start from 0
        
        # Read simulation data (skip header lines)
        df <- read_delim(f, delim = "\t", skip = 11, show_col_types = FALSE)
        df <- df %>%
            mutate(
                seed = seed,
                distribution = distribution,
                species_proportion = species_proportion,
                repetition = repetition
            )
        df
    })
}

# Load data from all simulation seeds
forceps_data <- map_dfr(seeds, read_forceps_file)
# Clean column names (remove # symbols)
colnames(forceps_data) <- gsub("#", "", colnames(forceps_data))
```
```

# Model Stochastic Variation Analysis

**Figure 2: Internal model variation across simulation seeds**

From previous replication studies, we know that averaging across 10 different seeds provides representative dynamics and reduces stochastic variation. Here we observe the substantial stochastic variation inherent in the model. The key question is whether variation between initialization methods remains detectable despite this stochastic noise.

```{r}
# Filter data for a single initialization method to examine stochastic variation
df <- forceps_data %>%
    filter(distribution == "mean", species_proportion == "fixe")

# Plot basal area trajectories for each species across different seeds
ggplot(df, aes(x = date, y = adultTreeBasalArea, color = factor(speciesId), group = interaction(seed, speciesId))) +
    geom_line() +
    theme_minimal() +
    labs(title = "Internal model variation (seeds)",
         x = "Date",
         y = "Basal Area (m²/ha)",
         color = "Species ID")
```

*Note: There appear to be initialization differences for species 17 and 14, which should be investigated as they should be identical across seeds.*


```{r}
# Comprehensive function to plot FORCEEPS simulation variations
# This function compares different initialization methods and their effects on forest dynamics
plot_forceps_variation <- function(data, variable, variation = "minmax", mode = "distrib") {
  library(ggplot2)
  library(dplyr)
  library(rlang)

  var_sym <- sym(variable)

  # Filter data based on analysis mode:
  # - "distribution": Compare diameter distribution methods (fixed species proportions)
  # - "species_proportion": Compare species allocation methods (fixed diameter distribution)
  filtered_data <- data %>%
    filter(
      (mode == "distribution" & species_proportion == "fixe") |
      (mode == "species_proportion" & distribution == "mean")
    ) %>%
    mutate(
      facet_label = if (mode == "distribution") paste0(distribution, "_", repetition)
                    else paste0(species_proportion, "_", repetition)
    )

  # Calculate summary statistics across simulation seeds
  summary_df <- filtered_data %>%
    filter(!is.na(!!var_sym)) %>%
    group_by(date, speciesId, facet_label) %>%
    summarise(
      mean_value = mean(!!var_sym, na.rm = TRUE),
      # Choose between min-max range or 95% confidence intervals
      lower = if (variation == "minmax") min(!!var_sym, na.rm = TRUE)
              else mean(!!var_sym, na.rm = TRUE) - 1.96 * sd(!!var_sym, na.rm = TRUE) / sqrt(n()),
      upper = if (variation == "minmax") max(!!var_sym, na.rm = TRUE)
              else mean(!!var_sym, na.rm = TRUE) + 1.96 * sd(!!var_sym, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  # Create visualization with ribbons showing variation and lines showing means
  ggplot(summary_df, aes(x = date, y = mean_value, color = factor(speciesId), fill = factor(speciesId))) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
    geom_line(size = 1) +
    facet_wrap(~facet_label) +
    theme_minimal() +
    labs(
      title = paste(variable, "par", mode, ifelse(variation == "minmax", "(ribbon = min–max)", "(IC 95%)")),
      x = "Date",
      y = variable,
      color = "Species ID",
      fill = "Species ID"
    )
}
```


# Effects of Diameter Distribution Methods

**Figure 3: Forest dynamics variation due to different diameter distribution initialization methods**

We fix species proportions to "fixe" (deterministic) and examine variation caused by different diameter distribution methods. The analysis covers multiple forest variables to assess the comprehensive impact of initialization choices.

```{r}
# Generate plots for key forest variables comparing diameter distribution methods
plot_forceps_variation(forceps_data, "adultTreeBasalArea_m2_ha", "minmax", mode = "distribution")
plot_forceps_variation(forceps_data, "adultTreeBiomass_t_ha", "minmax", mode = "distribution")
plot_forceps_variation(forceps_data, "deadBiomass_t_ha", "minmax", mode = "distribution")
plot_forceps_variation(forceps_data, "adultTreeNumber_ha", "minmax", mode = "distribution")
plot_forceps_variation(forceps_data, "saplingNumber_ha", "minmax", mode = "distribution")
```

# Effects of Species Proportion Methods

**Figure 4: Forest dynamics variation due to different species proportion initialization methods**

We fix diameter distribution to "mean" and examine variation caused by different species proportion allocation methods (deterministic vs. stochastic).

```{r}
# Generate plots for key forest variables comparing species proportion methods
plot_forceps_variation(forceps_data, "adultTreeBasalArea_m2_ha", "minmax", mode = "species_proportion")
plot_forceps_variation(forceps_data, "adultTreeBiomass_t_ha", "minmax", mode = "species_proportion")
plot_forceps_variation(forceps_data, "deadBiomass_t_ha", "minmax", mode = "species_proportion")
plot_forceps_variation(forceps_data, "adultTreeNumber_ha", "minmax", mode = "species_proportion")
plot_forceps_variation(forceps_data, "saplingNumber_ha", "minmax", mode = "species_proportion")
```

# Discussion and Future Directions

## Key Findings

This preliminary analysis reveals several important patterns in forest simulation initialization:

1. **Substantial stochastic variation**: The FORCEEPS model exhibits considerable internal variation across simulation seeds, emphasizing the need for multiple replications to obtain robust results.

2. **Initialization method effects**: Both diameter distribution and species proportion allocation methods produce detectable differences in forest dynamics, though these effects must be evaluated against the background of stochastic model variation.

3. **Variable sensitivity**: Different forest variables (basal area, biomass, tree numbers) show varying degrees of sensitivity to initialization choices.

## Methodological Implications

The analysis raises critical questions about inventory generation protocols:
- How should fractional tree representations be handled in small plots?
- What level of variation around initial data is appropriate given measurement uncertainties?
- How do initialization effects compare in magnitude to model stochastic variation?

## Future Research Priorities

To advance this work, future analyses should:

1. **Quantify initialization vs. stochastic variation**: Develop statistical methods to compare the magnitude of variation between initialization methods against internal model stochasticity.

2. **Test multiple forest types**: Expand analysis beyond the Retz forest to other forest ecosystems to assess the generalizability of findings.

3. **Develop standardized protocols**: Establish evidence-based guidelines for inventory generation that balance biological realism with computational efficiency.

4. **Long-term dynamics**: Assess whether initialization effects persist, amplify, or diminish over extended simulation periods.

This preliminary investigation provides a foundation for more comprehensive research into the critical but often overlooked question of how initial conditions affect forest simulation outcomes. 