---
title: "Retz Forest Data Visualization and Analysis"
author: "Clementine de Montgolfier"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r setup-options, include=FALSE}
# Configure knitr options for document generation
# Set working directory for all chunks to project root
knitr::opts_knit$set(root.dir = "../..")
# Configure chunk defaults: show code, suppress warnings and messages
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load-libraries, include=FALSE}
# Load required libraries for data analysis and visualization
library(tidyverse)    # Data manipulation and visualization
library(sf)           # Spatial data handling
library(tidyr)        # Data tidying
library(readxl)       # Excel file reading
library(patchwork)    # Combining multiple plots
library(knitr)        # Dynamic report generation
library(kableExtra)   # Enhanced table formatting
library(RColorBrewer) # Color palettes for visualization
```

# Data Preparation

Based on the assumptions made in Import_data.R
(particularly the selection of certain species)

```{r import-forest-data, include=FALSE}
# Load the processed forest data from RData file
load("data/retz/forest_data.RData")
# Extract species column names (columns 9 to 36 contain species data)
especes <- colnames(forest_data)[9:36]
```

```{r land-use-overview-plot, include=FALSE}
# Create a spatial visualization of land use and forest structure
# This plot shows the different forest management types and land occupation
ggplot(forest_data %>% 
       mutate(Structure.et.occupation.du.sol = factor(
         Structure.et.occupation.du.sol, 
         levels = c("F", "I", "R", "S", "T", "V", "Z", "M")))) +
  geom_sf(aes(geometry = geometry, fill = Structure.et.occupation.du.sol)) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),        # Remove grid lines
    axis.text.x = element_blank(),       # Remove x-axis labels
    axis.text.y = element_blank()) +     # Remove y-axis labels
  labs(title = "Land Use and Forest Structure", 
       fill = "Structure and Land Occupation") +
  # Define colors and labels for each land use category
  scale_fill_manual(
    values = c("#97e49d", "#0a6420", "#ff0000", "#28b128", 
               "#4f9446", "#a81616", "#abb2f7", "#ff15d8"), 
    labels = c("Regular high forest", "Irregular forest", "Non-forested (plantable)",
               "Coppice with standards", "Simple coppice", "Non-plantable",
               "Forested outside silviculture", "Service area"))
```

Data modifications applied:  
- Removal of plots without identifiers (n = 4)  
- Removal of non-forested plots (R non-forested plantable, V non-plantable, M service area) (n = 145), but kept forested areas outside silviculture (n=7)

```{r define-plotting-functions, include=FALSE}
# Define a utility function for creating combined spatial and histogram plots
map_plot <- function(data, fill_var) {
  # Create spatial map
  p1 <- ggplot(data) +
    geom_sf(aes(geometry = geometry, fill = !!sym(fill_var))) +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank()) +
    theme(legend.position = "none") +
    labs(title = fill_var)
  
  # Create frequency histogram
  p2 <- ggplot(data, aes(x = !!sym(fill_var))) +
    geom_bar(aes(fill = !!sym(fill_var), y = (..count..)/sum(..count..))) +
    theme_minimal() +
    # Rotate x-axis labels for better readability
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
    theme(legend.position = "none") +
    labs(x = "", y = "Number of plots")
  
  # Combine both plots
  p1 + p2
}
```

# Forest Stand Data Visualization (UEP)

## Average Basal Area

```{r basal-area-spatial-distribution, echo=FALSE, fig.width = 12, fig.height = 6}
# Create spatial map of average basal area
p1 <- ggplot(data = forest_data) +
      geom_sf(aes(geometry = geometry, fill = surf.terriere.moy)) +
      theme_minimal() +
      theme(
        panel.grid = element_blank(),        # Remove grid lines
        axis.text.x = element_blank(),       # Remove axis labels for cleaner map
        axis.text.y = element_blank()) +
      scale_fill_continuous(type = "viridis") + 
      labs(title = "Average Basal Area Distribution")

# Create density plot of basal area values
p2 <- ggplot(data = forest_data) +
      geom_density(aes(x = surf.terriere.moy, fill = surf.terriere.moy)) +
      theme_minimal()

# Combine spatial and statistical plots
p1 + p2
```

```{r basal-area-by-age, echo=FALSE, fig.width = 12, fig.height = 6}
# Analyze relationship between basal area and forest age
p1 <- ggplot(data = forest_data) +
  geom_boxplot(aes(x = surf.terriere.moy, y = median_age, 
                   group = median_age, color = surf.terriere.moy)) +
  geom_point(aes(x = surf.terriere.moy, y = median_age, color = surf.terriere.moy), 
             position = "jitter", alpha = 0.3) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Basal Area by Age Class",
       x = "Average Basal Area (m²/ha)",
       y = "Median Age (years)")

# Create density plot showing basal area distribution by age class
p2 <- ggplot(data = forest_data) +
    geom_density(aes(x = surf.terriere.moy, fill = surf.terriere.moy, 
                     color = factor(median_age))) +
    theme_minimal() +
    labs(color = "Age Class",
         x = "Average Basal Area (m²/ha)",
         y = "Density")

p1 + p2
```

## Dominant Species

```{r dominant-species-analysis, echo=FALSE, fig.width = 12, fig.height = 6}
# Use the utility function to create spatial and frequency plots for dominant species
map_plot(forest_data, "Essence.dominante.1")
```

## Stand Type Classification

```{r stand-type-visualization, echo=FALSE, fig.width = 12, fig.height = 6}
# Create a color palette for different stand types
palette_rainbow <- rainbow((length(unique(forest_data$Type.de.peuplement))))
type_peuplement_colors <- data.frame(palette_rainbow, sort(unique(forest_data$Type.de.peuplement)))
colnames(type_peuplement_colors) <- c("Color", "Type.de.peuplement")

# Create spatial map of stand types
p1 <- ggplot(forest_data %>% left_join(type_peuplement_colors)) +
  geom_sf(aes(geometry = geometry, fill = Type.de.peuplement)) +
  scale_fill_manual(values = type_peuplement_colors$Color, 
                    labels = type_peuplement_colors$Type.de.peuplement) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),        # Remove grid lines
    axis.text.x = element_blank(),       # Remove axis labels
    axis.text.y = element_blank()) +
  theme(legend.position = "none") +
  labs(title = "Stand Type Distribution")

# Create frequency bar chart of stand types
p2 <- ggplot(forest_data %>% left_join(type_peuplement_colors), 
             aes(x = Type.de.peuplement)) +
  geom_bar(aes(y = (..count..)/sum(..count..), fill = Type.de.peuplement)) +
  scale_fill_manual(values = type_peuplement_colors$Color, 
                    labels = type_peuplement_colors$Type.de.peuplement) +
  theme_minimal() +
  # Rotate x-axis labels for better readability
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.position = "none") +
  labs(x = "", y = "Number of plots")

p1 + p2
```

```{r dominant-stand-types-80-percent, echo=FALSE, fig.width = 12, fig.height = 6}
# Calculate stand types representing 80% of plots
forest_data %>%
  group_by(Type.de.peuplement) %>%
  summarise(Prop = n()/nrow(forest_data)) %>%
  arrange(-Prop) %>%
  mutate(Cum_Prop = cumsum(Prop)) %>%
  left_join(type_peuplement_colors) %>%
  filter(Cum_Prop <= 0.8) -> forest_data_80
```

There are `r nrow(forest_data_80)` stand types that represent 80% of all plots.

```{r plot-dominant-stand-types, echo=FALSE, fig.width = 6, fig.height = 6}
# Filter color palette to match the dominant stand types
type_peuplement_colors <- type_peuplement_colors %>%
  filter(Type.de.peuplement %in% forest_data_80$Type.de.peuplement)

# Create bar plot of dominant stand types
ggplot(forest_data_80, aes(x = Type.de.peuplement)) +
  geom_col(aes(y = Prop, fill = Type.de.peuplement)) +
  scale_fill_manual(values = type_peuplement_colors$Color, 
                    labels = type_peuplement_colors$Type.de.peuplement) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.position = "none") +
  labs(x = "", y = "Proportion of plots",
       title = "Most Common Stand Types (80% of plots)")
```

The stand types include:
- F: High forests (R,V non-forested)
- Species: HET (Beech), CHS (Sessile Oak), CHP (Pedunculate Oak), CHA (Hornbeam), P.L. (Corsican Pine), DOU (Douglas Fir), CHE (Native Oak)
- Various diameter classes
- The 4 most common (33%): Beech M, P, E, G diameter classes

## Dominant Diameter Category Class

```{r diameter-class-analysis, echo=FALSE, fig.width = 12, fig.height = 6}
# Analyze spatial and frequency distribution of diameter classes
map_plot(forest_data, "Classe.de.categorie.de.diametre.dominant")
```

```{r diameter-class-reference-table, echo=FALSE, fig.width = 12, fig.height = 6}
# Create reference table for diameter class codes and descriptions
table_ccd <- data.frame(
  CODE = c("S", "E", "IR", "1", "P", "M", "G", "G-", "G+", "T"),
  DESCRIPTION = c(
    "REGENERATION (H<3m)",
    "EDUCATION (non-commercial; d<10cm)",
    "IRREGULAR (no dominant diameter)",
    "POLES (10-15cm)",
    "SMALL TIMBER (20-25cm)",
    "MEDIUM TIMBER (30-45cm)",
    "LARGE TIMBER (50-65cm)",
    "Close to 50 cm",
    "Close to 65 cm",
    "VERY LARGE TIMBER (70cm+)"
  )
)

# Display the reference table
kable(table_ccd, col.names = c("CODE", "DESCRIPTION"), 
      caption = "Dominant Diameter Category Classes (CCD)")
```

## Coverage of Beech and Sessile Oak

Analysis of Sessile Oak combined with Pedunculate Oak, followed by Hornbeam

```{r species-coverage-maps, echo=FALSE, fig.width = 12, fig.height = 6}
# Create spatial map of Beech coverage
p1 <- ggplot(data = forest_data) +
    geom_sf(aes(geometry = geometry, fill = HET)) +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),      # Remove grid lines
        axis.text.x = element_blank(),     # Remove axis labels
        axis.text.y = element_blank()) +
    theme(legend.position = "none") +
    scale_fill_continuous(type = "viridis", na.value = "grey90") + 
    labs(title = "Beech Coverage")

# Create spatial map of combined Oak species coverage
p2 <- ggplot(data = forest_data %>% 
             rowwise() %>% 
             mutate(CH = sum(c(CHS, CHP), na.rm = TRUE)) %>% 
             ungroup()) +
    geom_sf(aes(geometry = geometry, fill = CH)) +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
    theme(legend.position = "none") +
    scale_fill_continuous(type = "viridis", na.value = "grey90") +
    labs(title = "Sessile and Pedunculate Oak")

# Create spatial map of Hornbeam coverage
p3 <- ggplot(data = forest_data) +
    geom_sf(aes(geometry = geometry, fill = CHA)) +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
    scale_fill_continuous(type = "viridis", na.value = "grey90") + 
    labs(title = "Hornbeam", fill = "% of forest cover")

# Arrange plots in a 2-column layout
p1 + p2 + p3 + plot_layout(ncol = 2)
```

# Comparison: Dominant Species vs Target Species

```{r species-objective-comparison, echo = FALSE, fig.width = 6}
# Create categorical variable comparing dominant and objective species
forest_data <- forest_data %>%
  mutate(essence_diff = case_when(
    Essence.dominante.1 == ess.obj.SIG ~ "Dominant species",
    Essence.dominante.2 == ess.obj.SIG ~ "Secondary species",
    TRUE ~ "Neither")) %>%
  mutate(essence_diff = factor(essence_diff, 
                              levels = c("Dominant species", "Secondary species", "Neither")))

# Create spatial visualization of species concordance
ggplot(forest_data) +
  geom_sf(aes(geometry = geometry, fill = essence_diff)) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank()) +
  scale_fill_manual(values = c("#44e26c", "#5a9aee", "#e46262"), 
                    labels = c("Dominant species", "Secondary species", "Neither")) +
  labs(title = "Concordance between Target and Dominant Species", 
       fill = "Concordance")

```

# Species Analysis: Dominance and Basal Area

## Dominant Species Analysis

```{r management-unit-aggregation, include=FALSE}
# Aggregate data by management unit to have one stand per management unit
# This creates weighted averages of species proportions based on area and basal area
ug_data <- forest_data %>%
  group_by(Identifiant.unite.de.gestion) %>%
  summarise(
    # Calculate weighted mean for all species columns
    across(all_of(especes), ~ weighted.mean(.x, surface.retenue*surf.terriere.moy, na.rm = TRUE)),
    # Calculate total surface area
    surface.tot = sum(surface.retenue),
    # Calculate weighted average basal area
    surf.terriere.moy.tot = sum(surface.retenue * surf.terriere.moy) / sum(surface.retenue)) %>%
  ungroup()

# Determine dominant species for each management unit
ug_data <- ug_data %>%
  mutate(
    essence.dominante.1 = apply(select(., all_of(especes)), 1, function(x) {
      if (all(is.na(x))) {
        NA_character_
      } else {
        especes[which.max(x)]
      }
    })
  )

# Extract unique dominant species
essence.dominante <- ug_data %>% 
  select(essence.dominante.1) %>% 
  unique() %>% 
  filter(!is.na(essence.dominante.1))
```

```{r dominant-species-comparison, echo=FALSE, fig.width = 12, fig.height = 4}
# Compare dominant species at management unit vs. individual stand level
p1 <- ggplot(ug_data, aes(x = essence.dominante.1)) +
  geom_bar(aes(y = (..count..), fill = essence.dominante.1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.position = "none") +
  labs(x = "", y = "Number of management units", 
       title = "Dominant Species by Management Unit")

p2 <- ggplot(forest_data, aes(x = Essence.dominante.1)) +
  geom_bar(aes(y = (..count..), fill = Essence.dominante.1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.position = "none") +
  labs(x = "", y = "Number of stands", 
       title = "Dominant Species by Individual Stand")

p1 + p2
```

There are `r nrow(essence.dominante)` species that are dominant in at least one management unit.
These are: `r paste(unique(essence.dominante$essence.dominante.1), collapse = ", ")`

## Basal Area Analysis

```{r basal-area-by-species, echo=FALSE, fig.width = 12, fig.height = 6}
# Calculate total basal area contribution by species across all forest stands
ug_surf.terriere <- forest_data %>%
  summarize(
    # Calculate weighted basal area for each species
    across(all_of(especes), ~ sum(.x * surf.terriere.moy*surface.retenue/100, na.rm = TRUE)),
  ) %>%
  # Reshape data from wide to long format
  pivot_longer(cols = all_of(especes), names_to = "Species", values_to = "Basal.Area") %>%
  mutate(Species = factor(Species, levels = especes)) %>%
  arrange(desc(Basal.Area)) %>%
  mutate(Species = factor(Species, levels = unique(Species)))

# Calculate total surface area for normalization
surf_tot = sum(forest_data$surface.retenue, na.rm = TRUE)

# Create bar plot of basal area by species
ggplot(ug_surf.terriere, aes(x = Species, y = Basal.Area/surf_tot)) +
  geom_col(aes(fill = Species)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.position = "none") +
  labs(x = "", y = "Basal Area (m²/ha)", title = "Basal Area Distribution by Species")

# Extract top 9 species by basal area for modeling
essence_surf.terriere <- ug_surf.terriere[1:9,] %>% 
  select(Species) %>% 
  unique()

# Calculate total basal area
st_tot = sum(ug_surf.terriere$Basal.Area)
st_tot/surf_tot
```

Proposed species to retain for modeling: `r paste(unique(essence_surf.terriere$Species), collapse = ", ")`

## Target Species Analysis

```{r target-species-analysis, echo=FALSE, fig.width = 10, fig.height = 6}
# Analyze frequency distribution of target species
forest_data %>% 
  select(ess.obj.SIG) %>% 
  table() %>% 
  as.data.frame() %>% 
  arrange(desc(Freq)) %>% 
  mutate(ess.obj.SIG = factor(ess.obj.SIG, levels = unique(ess.obj.SIG))) -> ess.obj.SIG

# Create bar plot of target species frequency
ggplot(ess.obj.SIG, aes(x = ess.obj.SIG)) +
  geom_col(aes(y = Freq, fill = ess.obj.SIG)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  theme(legend.position = "none") +
  labs(x = "", y = "Percentage of plots", title = "Target Species Distribution")

# Extract target species (excluding "SANS" = without target)
essence.objectif <- forest_data %>% 
  select(ess.obj.SIG) %>% 
  unique() %>% 
  filter(ess.obj.SIG!="SANS")
```

## FORCEPS Species Correspondence

```{r forceps-species-mapping, include=FALSE}
# Combine all relevant species from different analyses
essences <- unique(c(
  essence.dominante$essence.dominante.1,
  essence.objectif$ess.obj.SIG,
  essence_surf.terriere %>% pull(Species) %>% as.character()
)) %>% unlist()
essences <- table(essences) %>% as.data.frame()

# Load species correspondence table for FORCEPS model
forceps_species <- read.csv("data/corresponding_species.csv")
corresponding_species <- forceps_species %>%
  left_join(essences, by = c("Retz_Code" = "essences")) %>% 
  filter(!is.na(Freq))
```

```{r forceps-correspondence-table, echo = FALSE}
# Identify rows with missing FORCEPS codes
rows_to_highlight <- which(is.na(corresponding_species$Forceps_Code))

# Create table with highlighted rows for species without FORCEPS correspondence
corresponding_species  %>%
  kbl(caption = "Retz Species and FORCEPS Correspondence") %>%
  kable_classic(full_width = FALSE) %>%
  row_spec(rows_to_highlight, background = "#c2bd86")  # Highlight problematic species
```

**Species mapping notes:**
- Corsican Pine (Pin laricio) could be mapped to *Pinus nigra*
- Mixed birch species could be associated with *Betula pendula*

The remaining species have only one criterion (dominant species) and occur only in the plots shown below:

```{r rare-species-spatial-distribution, echo = FALSE}
# Define rare species codes
essence_prb <- c("AUL", "MER","PEU", "TIL", "EPS", "MEL")

# Create spatial map showing locations of rare dominant species
forest_data %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = Essence.dominante.1 %in% essence_prb)) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank()) +
  scale_fill_manual(values = c("#f5f5f5", "red"), labels = c("Common species", "Rare species")) +
  labs(title = "Spatial Distribution of Rare Dominant Species", 
       fill = "Species type")
```

```{r calculate-rare-species-statistics, echo = FALSE}
# Calculate statistics for rare species
surface_rare = forest_data %>% 
  filter(Essence.dominante.1 %in% essence_prb) %>% 
  pull(surface.retenue) %>% 
  sum()
placette_rare = forest_data %>% 
  filter(Essence.dominante.1 %in% essence_prb) %>% 
  nrow()
surface_tot = forest_data %>% 
  pull(surface.retenue) %>% 
  sum()
placette_tot = forest_data %>% nrow()
```

**Recommendation:** These rare species can be ignored initially as they are not representative of Retz Forest.

They represent `r placette_rare` plots (out of `r placette_tot`, or `r round(placette_rare/placette_tot*100, 2)`%) covering `r surface_rare` ha (out of `r surface_tot` ha, or `r round(surface_rare/surface_tot*100, 2)`%).

```{r rare-species-details-table, echo = FALSE}
# Create detailed table of stands with rare dominant species
forest_data %>%
  filter(Essence.dominante.1 %in% essence_prb) %>%
  select(Identifiant.peuplement.elementaire, Essence.dominante.1, Essence.dominante.2, 
         Classe.de.categorie.de.diametre.dominant, ess.obj, surface.retenue, surf.terriere.moy) %>%
  rename(
    "Stand ID" = Identifiant.peuplement.elementaire,
    "Dominant Species 1" = Essence.dominante.1,
    "Dominant Species 2" = Essence.dominante.2,
    "Diameter Class" = Classe.de.categorie.de.diametre.dominant,
    "Target Species" = ess.obj,
    "Area (ha)" = surface.retenue,
    "Mean Basal Area" = surf.terriere.moy
  ) %>%
  kbl(caption = "Stands with Rare Dominant Species") %>%
  kable_classic(full_width = FALSE)
```

# Aggregating Stands by Management Units

This approach may not be appropriate because stand sizes are larger than FORCEPS patches. Averaging would create artificial tree groups that don't exist in reality, ignoring plot heterogeneity. It might be better to retain only the majority surface of each plot.

```{r stand-size-distribution, echo = FALSE}
# Analyze distribution of retained surface areas
forest_data %>%
  filter(!is.na(surface.retenue)) %>%
  ggplot(aes(x = surface.retenue)) +
  geom_histogram(aes(y = ..density..), bins = 30) +
  geom_density(aes(y = ..density..), color = "blue") +
  theme_minimal() +
  # Add red line at 0.3 ha (maximum FORCEPS patch size)
  geom_vline(xintercept = 0.3, color = "red", linetype = "dashed") +
  labs(title = "Distribution of Retained Surface Area",
       x = "Retained Surface (ha)",
       y = "Density") +
  xlim(0,30)
```

The red line shows the maximum size of a FORCEPS patch (0.3 ha). For small plots, we might consider that both stand types coexist, but otherwise averaging across stands may not be relevant.

**Note:** There are probably some plots with very large retained surfaces - check for potential data errors.

# Available Water Capacity Analysis

```{r soil-water-capacity, echo = FALSE, fig.width = 10, fig.height = 6}
# Create spatial visualization of maximum available water capacity
ggplot(forest_data) +
  geom_sf(aes(geometry = geometry, fill = RUM)) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),    # Remove grid lines
    axis.text.x = element_blank(),   # Remove axis labels
    axis.text.y = element_blank()) +
  scale_fill_gradientn(colours = (brewer.pal(11, "Spectral"))) +
  labs(title = "Maximum Available Water Capacity", 
       fill = "RUM (mm)",
       caption = "Spatial distribution of soil water retention capacity across Retz Forest")
```